---
title: "Amazon sales"
subtitle: "Sales report"
author: "Sean Ng"
organization: "AIMdata"
date-modified: "18 May 2026"
execute: 
  echo: false
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE, 
                      message = FALSE, 
                      fig.width = 9)


library(tidyverse)
library(here)
library(janitor)
library(scales)
library(tidytext)
library(patchwork)
library(kableExtra)
library(viridis)
library(stringr)
library(ggh4x)
library(plotly)
library(ggrepel)
library(DT)
library(crosstalk)

`%out%` <- Negate(`%in%`)
options(scipen = 100)
theme_set(theme_light())
range_wna <- function(x){(x-min(x, na.rm = TRUE))/(max(x, na.rm = TRUE)-min(x, na.rm = TRUE))}

Mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

last_comma <- function(x, sep = ","){
  x <- sapply(strsplit(x, sep),"[")
  trimws(x[[length(x)]])
}
```

Fx rates come from here[https://www.exchangerates.org.uk/USD-INR-spot-exchange-rates-history-2023.html], average rate for 2023


```{r}
usd_to_inr <- 82.5738

amazon <- read_csv("./data/amazon.csv") |> 
  rename(discounted_price_inr = discounted_price, 
         actual_price_inr = actual_price) |> 
  mutate_at(
    vars(discounted_price_inr, 
         actual_price_inr), 
    ~ as.numeric(
      str_remove(str_remove(., "₹"), ",")
      )
  ) |> 
  mutate(discounted_price = discounted_price_inr / usd_to_inr, 
         actual_price = actual_price_inr / usd_to_inr) |> 
  mutate(discount_percentage = as.numeric(str_remove(discount_percentage, "%")), 
         discount_percentage = discount_percentage / 100)

categories <- amazon |> 
  distinct(product_id, category) |> 
  mutate(product_type = str_extract(category, "[^|]+$"), 
         first_category = str_extract(category, "^[^|]+")) |> 
  mutate(sub_category = str_split_i(category, "\\|", 2)) |>
  select(product_id, category = first_category, sub_category, product_type)

amazon <- amazon |> 
  select(-category) |> 
  left_join(
    categories, by = "product_id"
  )

```



## Product types 


```{r}
amazon |> 
  group_by(category, product_type) |> 
  summarise(rating_count = sum(rating_count, na.rm = TRUE), 
            rating = mean(rating, na.rm = TRUE), 
            .groups = "drop") |> 
  arrange(desc(rating_count)) |> 
  mutate(pc = rating_count / sum(rating_count)) |> 
  filter(rating_count > 100000) |> 
  ggplot(aes(x = rating_count, 
             y = fct_reorder(product_type, rating_count))) + 
  geom_col(aes(fill = category)) + 
  scale_fill_viridis_d() + 
  labs(x = "Number of ratings", 
       y = "", 
       title = "Most popular products on Amazon India", 
       fill = "Category") + 
  scale_x_continuous(labels = comma)
```

<br>



In-ear earphones are the most popular product type on Amazon India (given that you need to make a purchase before you can give a rating), followed by USB cables, smartphones and HDMI cables. 

Electronics, in general, seem to be the most popular category of products on Amazon India. It would be interesting to compare this other online sales platforms there. 

The plot below breaks down the most popular product types by sub-category. 

<br>


```{r fig.height=8, fig.width=9}
amazon |> 
  filter(sub_category %out% c("Tablets", "PowerAccessories", "Laptops", 
                              "Kitchen&Dining", "HomeMedicalSupplies&Equipment", 
                              "CarAccessories", "Arts&Crafts", "Monitors", 
                              "Microphones", "Electrical", "OfficeElectronics", 
                              "Components", "CraftMaterials")) |> 
  group_by(sub_category, product_type) |> 
  summarise(rating_count = sum(rating_count, na.rm = TRUE)) |> 
  group_by(sub_category) |> 
  slice_max(order_by = rating_count, n = 10, with_ties = FALSE) |> 
  mutate(product_type = str_sub(product_type, start = 1L, end = 20L)) |> 
  ggplot(aes(x = rating_count, 
             y = reorder_within(product_type, by = rating_count, within = sub_category))) + 
  geom_col(aes(fill = sub_category)) + 
  scale_fill_viridis_d(guide = "none") + 
  scale_x_continuous(labels = number_format(scale = 1/1000, suffix = "k")) +
  facet_wrap(~ sub_category, scales = "free") + 
  scale_y_reordered() + 
  theme(strip.background = element_rect(fill = "grey30"), 
        strip.text = element_text(size = 7), 
        axis.text.y = element_text(size = 5)) + 
  labs(title = "Most common product types by Amazon sub-category", 
       x = "Number of ratings", 
       y = "")
  
```

<br><br><br>



## Price and ratings

Amazon ratings, [similar](https://aimdata-labs.github.io/google_trip_reviews_kl_site/) to other online reviews on other platforms, are heavily-skewed towards the high-end. 

<br>

```{r warning=FALSE, message=FALSE}

amazon |> 
  # Filtering out product types with less than 1000 reviews 
  filter(product_type %out% c("AVReceivers&Amplifiers", "SATACables", "3DGlasses", 
                              "Earpads", "StovetopEspressoPots", "YogurtMakers", 
                              "StandMixers", "ElectricGrinders", "ColouredPaper", 
                              "RotiMakers", "InternalHardDrives", "TraditionalLaptops", 
                              "PhoneCharms", "TowerSpeakers", "StreamingClients")) |>
  group_by(product_id, rating) |>
  summarise(
    rating_count = sum(rating_count), 
    discounted_price = mean(discounted_price, na.rm = TRUE), 
    .groups = "drop") |> 
  mutate(rating_total = rating * rating_count, 
         price_total = discounted_price * rating_count) |> 
  group_by(product_id) |> 
  summarise(
    rating_total = sum(rating_total), 
    price_total = sum(price_total), 
    rating_count = sum(rating_count), 
    .groups = "drop"
  ) |> 
  mutate(mean_rating = rating_total / rating_count, 
         mean_discounted_price = price_total / rating_count) |> 
  left_join(
    amazon |> 
      distinct(product_id, product_name), 
    by = "product_id"
  ) |> 
  filter(rating_count >= 100) |>
  ggplot(aes(x = mean_rating, 
             y = mean_discounted_price)) + 
  geom_smooth(method = "lm", alpha = .5, 
              colour = "#c1121f") + 
  geom_point(aes(size = rating_count),
             colour = "#023047", 
             alpha = .4) + 
  scale_size_continuous(labels = comma, 
                        breaks = c(1000, 10000, 100000, 200000, 400000, 800000)) +
  scale_y_log10(breaks = c(0, 1, 3, 10, 30, 100, 300, 1000)) + 
  labs(x = "Mean rating", 
       y = "Mean discounted price (USD)", 
       title = "More expensive products seem to be rating more highly on Amazon India", 
       subtitle = "However, most ratings are between 4.0 and 4.5", 
       size = "Number of ratings") + 
  guides(size = guide_legend(override.aes = list(alpha = 1)))

```

<br>

```{r fig.height=6}
amazon |> 
  filter(
    product_type %out% (amazon |>
      count(product_type, sort = TRUE) |>
      filter(n == 1) |>
      pull(product_type))
  ) |> 
  group_by(category, product_type, rating) |>  
  summarise(
    rating_count = sum(rating_count, na.rm = TRUE),
    discounted_price = mean(discounted_price, na.rm = TRUE)
  ) |> 
  mutate(
    rating_total = rating * rating_count, 
    price_total = discounted_price * rating_count
  ) |> 
  group_by(category, product_type) |> 
  summarise(
    rating_total = sum(rating_total), 
    price_total = sum(price_total), 
    rating_count = sum(rating_count), 
    .groups = "drop"
  ) |>  
  mutate(mean_rating = rating_total / rating_count, 
         mean_discounted_price = price_total / rating_count) |> 
  mutate(label = ifelse(rating_count > 100000, product_type, "")) |> 
  ggplot(aes(x = mean_discounted_price, 
             y = mean_rating)) + 
  geom_smooth(method = "lm", 
              alpha = .5) + 
  geom_point(aes(colour = category, 
                 size = rating_count), 
             alpha = .7, 
             pch = 16) + 
  geom_text_repel(aes(label = label), 
                  size = 1.8) +
  scale_x_log10(breaks = c(1, 3, 10, 30, 100, 300)) + 
  scale_size_continuous(range = c(1, 10), 
                        labels = comma) +
  scale_colour_viridis_d(option = "turbo") + 
  labs(x = "Mean discounted price (USD)", 
       y = "Mean rating", 
       title = "Mean ratings and prices of Amazon products by category", 
       subtitle = "Tablets, power accessories and computer components are the most highly-reviewed", 
       colour = "Category", 
       size = "Number of ratings") + 
  guides(colour = guide_legend(override.aes = list(alpha = 1), 
                               order = 1))
  
```



<br>

Most product sub-categories demonstrate very little relationship between price and rating. 

For most product categories, higher prices do not always convert to higher product ratings

<br> 

```{r warning=FALSE, message=FALSE, fig.height=8}

backgrounds <- 
  list(
    element_rect(fill = "#780000"), 
    element_rect(fill = "#780000"),
    element_rect(fill = "#780000"),
    element_rect(fill = "#780000"),
    element_rect(fill = "#780000"),
    element_rect(fill = "#669bbc"),
    element_rect(fill = "#669bbc"),
    element_rect(fill = "#669bbc"),
    element_rect(fill = "#669bbc"),
    element_rect(fill = "#669bbc"),
    element_rect(fill = "#669bbc"),
    element_rect(fill = "#669bbc"),
    element_rect(fill = "#669bbc"),
    element_rect(fill = "#606c38"),
    element_rect(fill = "#606c38"),
    element_rect(fill = "#606c38"),
    element_rect(fill = "#606c38"),
    element_rect(fill = "#003049")
  )

category_list <- c(
"Accessories&Peripherals",
"Components",
"ExternalDevices&DataStorage",
"NetworkingDevices", 
"Printers,Inks&Accessories",

"Accessories",
"Cameras&Photography",
"GeneralPurposeBatteries&BatteryChargers",
"Headphones,Earbuds&Accessories",
"HomeAudio",
"HomeTheater,TV&Video",
"Mobiles&Accessories",
"WearableTechnology",

"CraftMaterials",
"Heating,Cooling&AirQuality",
"HomeStorage&Organization",
"Kitchen&HomeAppliances",

"OfficePaperProducts")

amazon |> 
  filter(!is.na(rating_count)) |> 
  group_by(category, sub_category, rating, discounted_price) |>
  summarise(rating_count = mean(rating_count), 
            .groups = "drop") |> 
  filter(sub_category %in% category_list) |>
  left_join(
    category_list |>
      as.data.frame() |>
      mutate(level = seq(1:18)), 
    by = c("sub_category" = "category_list")
  ) |> 
  mutate(sub_category = fct_reorder(sub_category, level)) |> 
  ggplot(aes(x = discounted_price, y = rating)) + 
  geom_point() + 
  geom_smooth() + 
  facet_wrap2(vars(sub_category), 
              scales = "free", 
              strip = strip_themed(
               background_x = backgrounds)) + 
  scale_x_log10() + 
  labs(title = "Amazon product ratings and prices", 
       subtitle = "Categories are: Computers & Accessories (red); Electronics (light blue); Office products (green); and Home & Kitchen (dark blue", 
       x = "Discounted price", 
       y = "Rating")
```

<br><br><br>



## Discounts

Almost all products on Amazon India are discounted. 

```{r}
amazon |> 
  filter(!is.na(rating)) |> 
  group_by(rating, discount_percentage) |> 
  summarise(count = n()) |> 
  ggplot(aes(x = discount_percentage, y = rating))  + 
  geom_smooth(method = "lm") +
  geom_point(aes(size = count), 
             alpha = .6, 
             pch = 16) + 
  scale_y_continuous(breaks = seq(2, 5, .5)) + 
  scale_x_continuous(labels = percent, breaks = seq(0, 1, .1)) +
  labs(x = "Discount percentage", 
       y = "Rating", 
       size = "Number\nof ratings", 
       title = "Customer ratings against percentage discount of Amazon items", 
       subitle = "Almost all items sold on Amazon India are discounted")
```

<br>

```{r}
amazon |> 
  filter(!is.na(rating)) |> 
  mutate(discount_group = 
           case_when(
             discount_percentage == 0 ~ "0%", 
             discount_percentage > 0 & discount_percentage <= .1 ~ "<=10%", 
             discount_percentage > .1 & discount_percentage <= .2 ~ "10.1%-20%", 
             discount_percentage > .2 & discount_percentage <= .3 ~ "20.1%-30%", 
             discount_percentage > .3 & discount_percentage <= .4 ~ "30.1%-40%", 
             discount_percentage > .4 & discount_percentage <= .5 ~ "40.1%-50%", 
             discount_percentage > .5 & discount_percentage <= .6 ~ "50.1%-60%", 
             discount_percentage > .6 & discount_percentage <= .7 ~ "60.1%-70%", 
             discount_percentage > .7 & discount_percentage <= .8 ~ "70.1%-80%", 
             discount_percentage > .8 & discount_percentage <= .9 ~ "80.1%-90%", 
             discount_percentage > .9 ~ ">90%")) |> 
  mutate(discount_group = fct_relevel(discount_group, 
                                      c("0%", "<=10%",
                                        "10.1%-20%",
                                        "20.1%-30%",
                                        "30.1%-40%",
                                        "40.1%-50%",
                                        "50.1%-60%",
                                        "60.1%-70%",
                                        "70.1%-80%",
                                        "80.1%-90%",
                                        ">90%"))) |> 
  group_by(discount_group) |> 
  summarise(median_rating = median(rating, na.rm = TRUE), 
            ratings = sum(rating_count, na.rm = TRUE)) |> 
  ggplot(aes(x = discount_group, y = ratings)) + 
  geom_col(aes(fill = median_rating)) +
  scale_fill_viridis() + 
  theme(axis.text.x = element_text(angle = 30, 
                                   hjust = .8)) + 
  scale_y_continuous(labels = number_format(scale = 1/1000, suffix = "k")) +
  labs(title = "Median rating by discount applied", 
       subtitle = "The highest median ratings come from products with no discounts", 
       x = "Discount applied", 
       y = "Number of ratings", 
       fill = "Number of\nratings")
```

<br>

If anything, the most highly-discounted items are less satisfactory to customers (having slightly lower ratings). It could also be that unsatisfactory and unpopular products have also been discounted so they would sell. 

Amongst non-discounted items (at 0% on the x-axis), there does seem to be a general range between 4 and 4.5 stars. 

What are the most commonly-discounted products


This code creates fewer objects in the environment, but it's still not really working well

So... crosstalk won't really work for histograms or freqpolies where the data is aggregated

Crosstalk doesn't seem to be working. Not sure why. 



```{r out.height="100%", out.height="100%", fig.width=7.8}

price_discount_categories_plot <- amazon |> 
  group_by(category, product_type) |> 
  summarise(
    mean_discount = mean(discount_percentage, na.rm = TRUE), 
    ratings_count = sum(rating_count, na.rm = TRUE), 
    mean_discounted_price = mean(discounted_price, na.rm = TRUE), 
    mean_actual_price = mean(actual_price, na.rm = TRUE), 
    .groups = "drop"
  ) |> 
  mutate(discount_label = round(mean_discount * 100, 2)) |> 
  ggplot(aes(x = mean_discounted_price, 
             y = mean_discount)) + 
  geom_smooth(method = "lm", alpha = .3) + 
  geom_point(aes(colour = category, 
                 text = paste0(
                   "Product type: ", product_type, ",", "\n", 
                   "Category: ", category, ",", "\n", 
                   "Mean actual price (USD): ", round(mean_actual_price, 2), ",", "\n",
                   "Mean discounted price (USD): ", round(mean_discounted_price, 2), ",", "\n",
                   "Mean discount: ", discount_label, "%"
                 ))) + 
  scale_x_log10(breaks = c(1, 3, 10, 30, 100, 300)) + 
  scale_y_continuous(labels = percent) +
  scale_colour_viridis_d(option = "turbo") + 
  labs(
    title = "Mean price and discount of product categories on Amazon India",
    x = "Mean discounted price in USD", 
    y = "Mean discount %", 
    colour = "Double-click to\nselect or reset"
  )

ggplotly(price_discount_categories_plot, tooltip = c("text")) |> 
  layout(height = 500, width = 800, 
         margin = list(t = 50)) |> 
  config(displayModeBar = FALSE)
```

<br>

Office products tend to have the lowest discounts, though many of these appear to be price-elastic low-cost staples which do not require additional incentives (such as having massive discounts) to attract sales. 

There is a tendency for more expensive items to have lower discounts, or perhaps, many heavily discounted items have dishonest "actual" prices: for instance, phone charms (the most discounted product type on Amazon India) have an average cost USD 12.10, but are offered at USD 1.20 at a discount of 90%. It is quite difficult to believe that a phone charm should cost USD 12.10, especially in India. 

We also see that the mean discount 


<br>

Most product sub-categories are largely defined by just a few major product types.

Amongst the most highly-rated (and most-purchased, since you need a purchase to make a rating) product types are USB calbes, HDMI cables, smartphones, MicroSDs, pen drives, smartwatches, wireless USB adapters and in-ear earbuds. 

<br><br><br>

## Review content


```{r}
amazon_bigrams <- amazon |> 
  select(review_id, review_content, user_id, category, sub_category, product_type) |>
  unnest_tokens(bigram, review_content, token = "ngrams", n = 2) |>  
  filter(!is.na(bigram)) |> 
  add_count(bigram) |>
  separate(bigram, c("word1", "word2"), sep = " ") |> 
  filter(!word1 %in% stop_words$word) |> 
  filter(!word2 %in% stop_words$word) |> 
  count(word1, word2, sub_category, sort = TRUE) |> 
  unite(bigram, word1, word2, sep = " ") |> 
  bind_tf_idf(bigram, sub_category, n)  
  
```



```{r fig.height = 9}
amazon_bigrams |> 
  filter(sub_category %out% c("Tablets", "PowerAccessories", "Laptops", 
                              "Kitchen&Dining", "HomeMedicalSupplies&Equipment", 
                              "CarAccessories", "Arts&Crafts", "Monitors", 
                              "Microphones", "Electrical", "OfficeElectronics", 
                              "Components", "CraftMaterials")) |> 
  filter(bigram %out% c("_sy88 jpg", "https m.media", "amazon.com images", 
                        "webp_402378 t2", "m.media amazon.com", 
                        "t2 images", "product https")) |> 
  group_by(sub_category) |> 
  slice_max(order_by = n, n = 10, with_ties = FALSE) |> 
  ggplot(aes(x = n, y = reorder_within(bigram, by = n, within = sub_category))) + 
  geom_col(aes(fill = sub_category)) +
  facet_wrap(~ sub_category, scales = "free") + 
  scale_y_reordered() + 
  scale_fill_viridis_d(guide = "none") + 
  labs(title = "Most common bigrams by product sub-category", 
       x = "Count of bigrams within sub-category", 
       y = "") + 
  theme(strip.background = element_rect(fill = "grey20"), 
        strip.text = element_text(size = 7))
```

<br>

Given the enshittified state of consumer products, do people have an accurate idea of what quality is? 

Maybe we should try and re-scale the review scores

The meaninglessness of the word "quality". 


