---
title: "Amazon sales"
subtitle: "Sales report"
author: "Sean Ng"
organization: "AIMdata"
date: '`r format(Sys.Date())`'
output: html_document: 
  self_contained: TRUE
params: 
  dataset: NULL
execute: 
  echo: false
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE, 
                      message = FALSE, 
                      fig.width = 9)


library(tidyverse)
library(janitor)
library(scales)
library(patchwork)
library(viridis)
library(stringr)
library(broom)
library(ggh4x)
library(plotly)
library(ggrepel)


`%out%` <- Negate(`%in%`)
options(scipen = 100)
theme_set(theme_light())
range_wna <- function(x){(x-min(x, na.rm = TRUE))/(max(x, na.rm = TRUE)-min(x, na.rm = TRUE))}

Mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

last_comma <- function(x, sep = ","){
  x <- sapply(strsplit(x, sep),"[")
  trimws(x[[length(x)]])
}
```

## Introduction 

This is a summary report of [sales data](https://www.kaggle.com/datasets/karkavelrajaj/amazon-sales-dataset) from Amazon India compiled by [Karkavelaraja J.](https://www.linkedin.com/in/karkavelraja-j-a758031b0/) 

The dataset contains more than 1,000 Amazon product ratings and reviews from 2023. 

<br><br><br>

```{r}
# Fx rates come from here[https://www.exchangerates.org.uk/USD-INR-spot-exchange-rates-history-2023.html], the average rate for 2023 has been used. 

usd_to_inr <- 82.5738

amazon <- read_csv("./data/amazon.csv") |> 
  rename(discounted_price_inr = discounted_price, 
         actual_price_inr = actual_price) |> 
  mutate_at(
    vars(discounted_price_inr, 
         actual_price_inr), 
    ~ as.numeric(
      str_remove(str_remove(., "₹"), ",")
      )
  ) |> 
  mutate(discounted_price = discounted_price_inr / usd_to_inr, 
         actual_price = actual_price_inr / usd_to_inr) |> 
  mutate(discount_percentage = as.numeric(str_remove(discount_percentage, "%")), 
         discount_percentage = discount_percentage / 100)

categories <- amazon |> 
  distinct(product_id, category) |> 
  mutate(product_type = str_extract(category, "[^|]+$"), 
         first_category = str_extract(category, "^[^|]+")) |> 
  mutate(sub_category = str_split_i(category, "\\|", 2)) |>
  select(product_id, category = first_category, sub_category, product_type)

amazon <- amazon |> 
  select(-category) |> 
  left_join(
    categories, by = "product_id"
  )
```


## Product types 

The plot below shows the most-rated (and antecedently, the most-purchased, given that all ratings require a purchase) product types on Amazon India. In-ear earphones are the most popular product type on Amazon India, followed by USB cables, smartphones and HDMI cables. 

<br> 

```{r}
amazon |> 
  group_by(category, product_type) |> 
  summarise(rating_count = sum(rating_count, na.rm = TRUE), 
            rating = mean(rating, na.rm = TRUE), 
            .groups = "drop") |> 
  arrange(desc(rating_count)) |> 
  mutate(pc = rating_count / sum(rating_count)) |> 
  filter(rating_count > 100000) |> 
  ggplot(aes(x = rating_count, 
             y = fct_reorder(product_type, rating_count))) + 
  geom_col(aes(fill = category)) + 
  scale_fill_viridis_d() + 
  labs(x = "Number of ratings", 
       y = "", 
       title = "Most popular products on Amazon India", 
       fill = "Category") + 
  scale_x_continuous(labels = comma)
```

<br>

Electronics, in general, seem to be the most popular category of products on Amazon India. This post-modern age does seem to require and generate an excess of electronic detritus. It would be interesting to compare this other online sales platforms there. 

The plot below provides additional detail and shows the 10 most popular product types by sub-category. As is expected, many sub-categories are dominated by one or two large product types, with a "long tail" of more miscellaneous options. 

<br>


```{r fig.height=8, fig.width=9}
amazon |> 
  filter(sub_category %out% c("Tablets", "PowerAccessories", "Laptops", 
                              "Kitchen&Dining", "HomeMedicalSupplies&Equipment", 
                              "CarAccessories", "Arts&Crafts", "Monitors", 
                              "Microphones", "Electrical", "OfficeElectronics", 
                              "Components", "CraftMaterials")) |> 
  group_by(sub_category, product_type) |> 
  summarise(rating_count = sum(rating_count, na.rm = TRUE)) |> 
  group_by(sub_category) |> 
  slice_max(order_by = rating_count, n = 10, with_ties = FALSE) |> 
  mutate(product_type = str_sub(product_type, start = 1L, end = 20L)) |> 
  ggplot(aes(x = rating_count, 
             y = reorder_within(product_type, by = rating_count, within = sub_category))) + 
  geom_col(aes(fill = sub_category)) + 
  scale_fill_viridis_d(guide = "none") + 
  scale_x_continuous(labels = number_format(scale = 1/1000, suffix = "k")) +
  facet_wrap(~ sub_category, scales = "free") + 
  scale_y_reordered() + 
  theme(strip.background = element_rect(fill = "grey30"), 
        strip.text = element_text(size = 7), 
        axis.text.y = element_text(size = 5)) + 
  labs(title = "Most common product types by Amazon sub-category", 
       x = "Number of ratings", 
       y = "")
  
```

<br><br><br>

## Price and ratings

Amazon ratings, [similar](https://aimdata-labs.github.io/google_trip_reviews_kl_site/) to other online reviews on other platforms, are heavily-skewed towards the high-end. 

<br>

```{r warning=FALSE, message=FALSE}

amazon |> 
  # Filtering out product types with less than 1000 reviews 
  filter(product_type %out% c("AVReceivers&Amplifiers", "SATACables", "3DGlasses", 
                              "Earpads", "StovetopEspressoPots", "YogurtMakers", 
                              "StandMixers", "ElectricGrinders", "ColouredPaper", 
                              "RotiMakers", "InternalHardDrives", "TraditionalLaptops", 
                              "PhoneCharms", "TowerSpeakers", "StreamingClients")) |>
  group_by(product_id, rating) |>
  summarise(
    rating_count = sum(rating_count), 
    discounted_price = mean(discounted_price, na.rm = TRUE), 
    .groups = "drop") |> 
  mutate(rating_total = rating * rating_count, 
         price_total = discounted_price * rating_count) |> 
  group_by(product_id) |> 
  summarise(
    rating_total = sum(rating_total), 
    price_total = sum(price_total), 
    rating_count = sum(rating_count), 
    .groups = "drop"
  ) |> 
  mutate(mean_rating = rating_total / rating_count, 
         mean_discounted_price = price_total / rating_count) |> 
  left_join(
    amazon |> 
      distinct(product_id, product_name), 
    by = "product_id"
  ) |> 
  filter(rating_count >= 100) |>
  ggplot(aes(x = mean_rating, 
             y = mean_discounted_price)) + 
  geom_smooth(method = "lm", alpha = .5, 
              colour = "#c1121f") + 
  geom_point(aes(size = rating_count),
             colour = "#023047", 
             alpha = .4) + 
  scale_size_continuous(labels = comma, 
                        breaks = c(1000, 10000, 100000, 200000, 400000, 800000)) +
  scale_y_log10(breaks = c(0, 1, 3, 10, 30, 100, 300, 1000)) + 
  labs(x = "Mean rating", 
       y = "Mean discounted price (USD)", 
       title = "More expensive products seem to be rating more highly on Amazon India", 
       subtitle = "However, most ratings are between 4.0 and 4.5", 
       size = "Number of ratings") + 
  guides(size = guide_legend(override.aes = list(alpha = 1)))

```

<br>

Most product sub-categories demonstrate very little relationship between price and rating, or, higher prices do not always translate to higher product ratings. 

<br>

```{r warning=FALSE, fig.height=8}

backgrounds <- 
  list(
    element_rect(fill = "#780000"), 
    element_rect(fill = "#780000"),
    element_rect(fill = "#669bbc"),
    element_rect(fill = "#669bbc"),
    element_rect(fill = "#669bbc"),
    element_rect(fill = "#669bbc"),
    element_rect(fill = "#669bbc"),
    element_rect(fill = "#669bbc"),
    element_rect(fill = "#780000"),
    element_rect(fill = "#669bbc"),
    element_rect(fill = "#606c38"),
    element_rect(fill = "#003049"),
    element_rect(fill = "#669bbc"),
    element_rect(fill = "#780000"),
    element_rect(fill = "#780000"),
    element_rect(fill = "#003049"),
    element_rect(fill = "#003049"),
    element_rect(fill = "#003049")
  )

amazon |> 
  filter(!is.na(rating_count)) |> 
  group_by(category, sub_category, rating, discounted_price) |>
  summarise(rating_count = mean(rating_count), 
            .groups = "drop") |> 
  filter(sub_category %in% c("Accessories", "Accessories&Peripherals", 
                              "Cameras&Photography", "Components", 
                              "CraftMaterials",
                              "ExternalDevices&DataStorage", 
                              "GeneralPurposeBatteries&BatteryChargers", 
                              "Headphones,Earbuds&Accessories",
                              "Heating,Cooling&AirQuality", 
                              "HomeAudio", "HomeStorage&Organization", 
                              "HomeTheater,TV&Video", 
                              "Kitchen&HomeAppliances", 
                              "Mobiles&Accessories", 
                              "NetworkingDevices", "OfficePaperProducts", 
                              "Printers,Inks&Accessories", 
                              "WearableTechnology")) |> 
  ggplot(aes(x = discounted_price, y = rating)) + 
  geom_point() + 
  geom_smooth() + 
  facet_wrap2(vars(sub_category), 
              scales = "free", 
              strip = strip_themed(
               background_x = backgrounds)) + 
  scale_x_log10() + 
  labs(title = "Amazon product ratings and prices", 
       subtitle = "Categories are: Computers & Accessories; Electronics; Office products; and Home & Kitchen")
```

<br><br><br>

## Discounts

"Discounts" are pervasive part of online (and offline) marketing. We see from the plot below that almost all products on Amazon India are discounted, displaying in full the *cheapness* of this promotional strategy. 

From the plot below, it does, however, seem that heavily-discounted items were less satisfactory to consumers. It could also be that low-quality and unpopular products were discounted so that they would sell. 

<br>

```{r}
amazon |> 
  filter(!is.na(rating)) |> 
  mutate(discount_group = 
           case_when(
             discount_percentage == 0 ~ "0%", 
             discount_percentage > 0 & discount_percentage <= .1 ~ "<=10%", 
             discount_percentage > .1 & discount_percentage <= .2 ~ "10.1%-20%", 
             discount_percentage > .2 & discount_percentage <= .3 ~ "20.1%-30%", 
             discount_percentage > .3 & discount_percentage <= .4 ~ "30.1%-40%", 
             discount_percentage > .4 & discount_percentage <= .5 ~ "40.1%-50%", 
             discount_percentage > .5 & discount_percentage <= .6 ~ "50.1%-60%", 
             discount_percentage > .6 & discount_percentage <= .7 ~ "60.1%-70%", 
             discount_percentage > .7 & discount_percentage <= .8 ~ "70.1%-80%", 
             discount_percentage > .8 & discount_percentage <= .9 ~ "80.1%-90%", 
             discount_percentage > .9 ~ ">90%")) |> 
  mutate(discount_group = fct_relevel(discount_group, 
                                      c("0%", "<=10%",
                                        "10.1%-20%",
                                        "20.1%-30%",
                                        "30.1%-40%",
                                        "40.1%-50%",
                                        "50.1%-60%",
                                        "60.1%-70%",
                                        "70.1%-80%",
                                        "80.1%-90%",
                                        ">90%"))) |> 
  group_by(discount_group) |> 
  summarise(median_rating = median(rating, na.rm = TRUE), 
            ratings = sum(rating_count, na.rm = TRUE)) |> 
  ggplot(aes(x = discount_group, y = ratings)) + 
  geom_col(aes(fill = median_rating)) +
  scale_fill_viridis() + 
  theme(axis.text.x = element_text(angle = 30, 
                                   hjust = .8)) + 
  scale_y_continuous(labels = number_format(scale = 1/1000, suffix = "k")) +
  labs(title = "Median rating by discount applied", 
       subtitle = "The highest median ratings come from products with no discounts", 
       x = "Discount applied", 
       y = "Number of ratings", 
       fill = "Median\nrating")
```

<br>

We can investigate this phenomenon in a bit more detail using the interactive plot below. 

<br>

```{r out.height="100%", out.height="100%", fig.width=7.8}

price_discount_categories_plot <- amazon |> 
  group_by(category, product_type) |> 
  summarise(
    mean_discount = mean(discount_percentage, na.rm = TRUE), 
    ratings_count = sum(rating_count, na.rm = TRUE), 
    mean_discounted_price = mean(discounted_price, na.rm = TRUE), 
    mean_actual_price = mean(actual_price, na.rm = TRUE), 
    .groups = "drop"
  ) |> 
  mutate(discount_label = round(mean_discount * 100, 2)) |> 
  ggplot(aes(x = mean_discounted_price, 
             y = mean_discount)) + 
  geom_smooth(method = "lm", alpha = .3) + 
  geom_point(aes(colour = category, 
                 text = paste0(
                   "Product type: ", product_type, ",", "\n", 
                   "Category: ", category, ",", "\n", 
                   "Mean actual price (USD): ", round(mean_actual_price, 2), ",", "\n",
                   "Mean discounted price (USD): ", round(mean_discounted_price, 2), ",", "\n",
                   "Mean discount: ", discount_label, "%"
                 ))) + 
  scale_x_log10(breaks = c(1, 3, 10, 30, 100, 300)) + 
  scale_y_continuous(labels = percent) +
  scale_colour_viridis_d(option = "turbo") + 
  labs(
    title = "Mean price and discount of product categories on Amazon India",
    x = "Mean discounted price in USD", 
    y = "Mean discount %", 
    colour = "Double-click to\nselect or reset"
  )

ggplotly(price_discount_categories_plot, tooltip = c("text")) |> 
  layout(height = 500, width = 800, 
         margin = list(t = 50)) |> 
  config(displayModeBar = FALSE)
```

<br>

Office products tend to have the lowest discounts, though many of these appear to be price-elastic low-cost staples which do not require additional incentives (such as having massive discounts) to attract sales. 

There is a tendency for more expensive items to have lower discounts. Or perhaps, it is that many heavily "discounted" items have dishonest "actual" prices: for instance, phone charms (the most discounted product type on Amazon India) have an average "actual" cost of USD 12.10. This is a nonsensical price, but it does allow them to be offered at the average discounted price of USD 1.20, with a 90% discount. 

<br><br><br>

## Review content


```{r}
amazon_bigrams <- amazon |> 
  select(review_id, review_content, user_id, category, sub_category, product_type) |>
  unnest_tokens(bigram, review_content, token = "ngrams", n = 2) |>  
  filter(!is.na(bigram)) |> 
  add_count(bigram) |>
  separate(bigram, c("word1", "word2"), sep = " ") |> 
  filter(!word1 %in% stop_words$word) |> 
  filter(!word2 %in% stop_words$word) |> 
  count(word1, word2, sub_category, sort = TRUE) |> 
  unite(bigram, word1, word2, sep = " ") |> 
  bind_tf_idf(bigram, sub_category, n)  
  
```

Let us now take a look at the most common bigrams (two words) per product sub-category. 

These bigrams provide insight into what consumers believe to be the most relevant considerations when purchasing a type of product. 

For instance, in the Headphones and Earbuds sub-category, we see the that the top two bigrams are "sound quality" and "noise cancellation", followed by "1 hour" and "minutes approx", which are probably references to the charging speed of wireless earbuds. 



<br>

```{r fig.height = 9}
amazon_bigrams |> 
  filter(sub_category %out% c("Tablets", "PowerAccessories", "Laptops", 
                              "Kitchen&Dining", "HomeMedicalSupplies&Equipment", 
                              "CarAccessories", "Arts&Crafts", "Monitors", 
                              "Microphones", "Electrical", "OfficeElectronics", 
                              "Components", "CraftMaterials")) |> 
  filter(bigram %out% c("_sy88 jpg", "https m.media", "amazon.com images", 
                        "webp_402378 t2", "m.media amazon.com", 
                        "t2 images", "product https")) |> 
  group_by(sub_category) |> 
  slice_max(order_by = n, n = 10, with_ties = FALSE) |> 
  ggplot(aes(x = n, y = reorder_within(bigram, by = n, within = sub_category))) + 
  geom_col(aes(fill = sub_category)) +
  facet_wrap(~ sub_category, scales = "free") + 
  scale_y_reordered() + 
  scale_fill_viridis_d(guide = "none") + 
  labs(title = "Most common bigrams by product sub-category", 
       x = "Count of bigrams within sub-category", 
       y = "") + 
  theme(strip.background = element_rect(fill = "grey20"), 
        strip.text = element_text(size = 7))
```

<br> 

